Automated Formal Synthesis\\ of Safe Digital Controllers for Continuous Plants

### 2018-12-20 automatica/paper.tex
Removing commented out text
in the event that we want it back, we can use git history

%All we need to do is define the nature of $\mathcal{A}X_0$.
%Let $\mat{S}\mat{J}\mat{S}^{-1}=\mat{A}$ be the eigen-decomposition of
%$\mat{A}$, where $\mat{S}$ is the set of generalised eigenvectors of
%$\mat{A}$ and $\mat{J}$ is the Jordan Form containing the eigenvalues of
%$\mat{A}$ in its main diagonal.  We correspondingly define
%$\mathcal{J}=\mat{S}^{-1}\mathcal{A}\mat{S}$, which is such that
%$\mathcal{J} \supseteq \bigcup_{k \in [0\ \infty)} \mat{J}^k$.
%Since abstract acceleration operates in the eigenspace, wemust deal with some preliminary transformations.
%Given a set $X=\{ \vec{x} \mid \mat{C}\vec{x} \leq \vec{d}\}$,
%the linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to the set 
%\begin{equation}
%X'= \mat{S}^{-1}X = \{\vec{x}' \mid \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
%\label{eq:eigenInit}
%\end{equation}
%Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
%${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
%abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
%transform equation \eqref{eq:aa_reachtube} into: 
%\begin{equation}
%\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', 
%\end{equation}
%where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
%polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
%$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
%tableau defined by $\mathcal{J}$ and objective functions derived from each
%vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
%\ref{alg:findReachTube} below).  We note, however, that if we wish to use
%numerical analysis in the decomposition of $\mat{A}$, the Tableau in
%$\mathcal{J}$ contains small intervals given by the errors, which cannot be
%processed efficiently by a regular simplex and require a new procedure.
%The calculation of the abstract reach tube via abstract acceleration is
%encompassed in Algorithm~\ref{alg:findReachTube}, which can be summarised as
%follows (its steps are denoted in parentheses):
%\item We perform unsound eigen-decomposition using an existing algebra package (lines \ref{alg:calcEigenvalues}-\ref{alg:calcEigenvectors}).
%\item Then we restore soundness to the results using the methods described in \cite{cattaruzza2017sound} (lines \ref{alg:soundEigenvalues}-\ref{alg:soundEigenvectors}).
%\item The inverse of the matrix of eigenvectors is calculated after soundness is restored, using interval arithmetic in order to ensure its soundness (line \ref{alg:inverse}).
%\item The abstract dynamics are obtained by evaluating the convex hull of all powers of eigenvalues up to the desired number of iterations as described in \cite{cattaruzza2015unbounded} (line \ref{alg:aa}). 
%\item Using Equation \eqref{eq:eigenInit}, transform the initial state into the eigenspace (line \ref{alg:eigeninit}).
%\item Extract the vertices of the eigen-polyhedron $X_0'$ using the algorithms described in \cite{cattaruzza2015unbounded,cattaruzza2017sound} (line \ref{alg:getVertices}). 
%\item Calculate the mapped \emph{abstract reach tube} ${\hat{X}'^\sharp}$
%that over-approximates the image of the reach tube in the eigenspace (line
%\ref{alg:getReachTube}).  This is achieved by evaluating a set of objective
%functions using the abstract dynamics as a simplex Tableau, and via a sound
%simplex described in \cite{cattaruzza2017sound}.  The objective functions are
%defined by the vertices of $X'_0$, denote as $V_0$, and by the desired
%template directions, such that
%\begin{equation*}
%\vec{w}_{ij}= \vec{v}_i \circ \vec{t}_j \text{ with } \vec{v}_i \in V_0 \text{ and } \vec{t}_j \in T,
%\end{equation*}
%where $T$ is the set of template directions, and $\circ$ denotes a
%component-wise multiplication yielding a vector.
%\item Using Equation~\eqref{eq:eigenInit}, find the reach tube
%$\hat{X}^\sharp=\mat{S}{\hat{X}'^\sharp}$ (line \ref{alg:getNormReachTube}).
%\begin{algorithm}[]
%\small
%\begin{algorithmic}[1]
%\Function{$\mathit{findAbstractReachTube}()$}{}
%\State $\hat{\mat{J}}=$ calculateEigenvalues($ \mat{A}$)
%\label{alg:calcEigenvalues}
%\State $\hat{\mat{S}}=$ calculateEigenvectors($ \mat{A}$)
%\label{alg:calcEigenvectors}
%\State $[\mat{J}]=$ soundifyEigenvalues($\mat{A},\mat{J},\mat{S}$)
%\label{alg:soundEigenvalues}
%\State $[\mat{S}]=$ soundifyEigenvectors($\mat{A},[\mat{J}],\mat{S}$)
%\label{alg:soundEigenvectors}
%\State $[\mat{S}]^{-1}=$ calculateInverse($[\mat{S}]$)
%\label{alg:inverse}
%\State $\mathcal{J}=getAbstractDynamics([\mat{J}])$
%\label{alg:aa}
%\State $[X_0']=$ transformInitialSpace($X_0,\mat{S}$)
%\label{alg:eigeninit}
%\State $[V_0]=$ getVertices($[X_0']$)
%\label{alg:getVertices}
%\State ${\hat{X}'^\sharp}=$ getReachTube($\mathcal{J},[V_0]$)
%\label{alg:getReachTube}
%\State ${\hat{X}^\sharp}=$ transformReachTube(${\hat{X}'^\sharp},[\mat{S}]^{-1}$)
%\label{alg:getNormReachTube}
%\EndFunction
%\end{algorithmic}
%\caption{Calculation of Abstract Reach Tube Using Abstract Acceleration}
%\label{alg:findReachTube}
%\end{algorithm}

%%-------------------------------
%\subsection{Overapproximating continuous time dynamics using discrete models}
%\label{sec:cont_aa}
%%-------------------------------
%
%Thus far we wave discussed the Abstract Accelertion of Discrete Time
%systems.  However, our plant models derive from the discretization of
%continuous time systems.  We therefore seek to establish a sound
%overapproximation which does not only encompass the selected discretization,
%but any chosen discretization in the time domain.
%
%A dynamical system is a system in which a function describes the progression
%of a state over time.  In a continuous domain with linear dynamics, it is
%described by a first order Ordinary Differential Equation (ODE).
%%
%\begin{equation}
%\dot{x}(t)=\mat{A}_o\vec{x}(t)+\mat{B}_o\vec{u}(t)
%\label{eq:dynamical}
%\end{equation}
%%
%Furthermore, a control system may have a derived
%output that is a linear combination of its states and inputs, which may
%restricts the observability of the statespace from the output space.
%%
%\begin{equation}
%\vec{y}(t)=\mat{C}_o\vec{x}(t)+\mat{D}_o\vec{u}(t)
%\end{equation}
%%
%Discretization of a continuous dynamical system turns the ODE into a
%difference equation, assuming zero-order hold for the input $\vec{u}$, to
%%
%\begin{align}
%\label{eq:discretization}
%\vec{x}_{k+1} &= \mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
%y_k &= \mat{C}_d \vec{x}_ k + \mat{D}_d \vec{u}_ k 
%\end{align}
%where
%\begin{align}
%\label{eq:discretize}
%\mat{A}_d &= e^{\mat{A}_o T_s} = \mathcal{L}^{-1} { ( s \mat{I} - \mat{A}_o )^{-1} }_{t = T_s}\\
%\mat{B}_d &= \int_{t = 0}^{T_s} e^{\mat{A}_o t} dt\ \mat{B}_o = \mat{A}_o^{-1} ( \mat{A}_d - \mat{I} ) \mat{B}_o\\
%\mat{C}_d &= \mat{C}_o\\
%\mat{D}_d &= \mat{D}_o \\
%\end{align}
%and $T_s$ is the sample time. Then
%\begin{align*}
%x(kT)=x_k \wedge y(kT) = y_k, \forall k
%\end{align*}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%\subsubsection{Acceleration of continuous-time models}
%\label{sec:continuous_time_accel}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The time discretization in Sec.~\ref{sec:model} only allows us to evaluate
the state of the plant at integer multiples of the sample time.  
%\note{Unlike the CAV18 paper, these gain matrices have not been discussed yet. }
However, in order to synthesize a soundly safe controller, 
% $\mat{K}$, 
we must be able to ensure that the safety of the plant holds at any continuous point in time.  
In order to do so, we accelerate the dynamics of the plant according to the continuous-time model
semantics in Sec.~\ref{sec:model}. 
% and compute the state $\vec{x}_T$ at a given time $T$,
%%
%% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
%%
%namely the state $\vec{x}(t)$ of the model 
%%$\dot{\vec{x}}(t)=\mat{A}
%at time $t=T$.

%\textcolor{red}{[The following three subsections take too much space, plus the results are mostly known or easily derivable with simple algebra: we need to shorten and simplify it. Dario: can you give it a crack?]}
%\textcolor{blue}{[I merged the first two]}

%\note{ [add?] In this section we work with fully observed models, namely where $y = x$. } 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Accelerated\! dynamics\! for\! continuous-time\! models\! with\! parametric\! inputs}\label{sec:real_discrete_param_inputs}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
% 
Given a specific time $T$, the state $\vec{x}(t=T)$ can be expressed as a
linear function of initial state $\vec{x}_0$ and parametric input $\vec{u}$ as follows.

@@

%Abstract acceleration works by finding support functions of the set that encompasses the matrices $\mat{A}^k$.
%Support functions are scalars that measure the distance of a hyperplane tangent to the set boundary to the origin.
%In the case of continuous spaces, this translates to calculating the derivatives of the envelope in order to find the direction of the support vector (normal to the hyperplane), and calculating the distance to the corresponding point where the hyperplane touches the set (at which the derivative was calculated).
%The direction of the vector is determined by evaluating the support function of another point in the curve,
%which is known to be smaller than that of the selected point, and the chosen hyperplanes are aligned with the eigenvectors, hence we only look at eigenvalue pairs to calculate the derivatives~\cite{cattaruzza2015unbounded}.
%\subsection{State-space representation of physical systems} 
%\label{ssec:ssrepresentation}
%We consider models of physical plants expressed as ordinary differential
%equations (ODEs), which we assume to be controllable and under full state
%information (i.e., we have access to all the model variables):
%\begin{align}
%\label{eq:ode}
%\dot{x}(t) = Ax(t)+ B u(t), \quad x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, \\ \nonumber A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}, 
%\end{align}
%where $t \in \mathbb R_0^+$, where $A$ and $B$ are matrices that fully
%specify the continuous plant, and with initial states set as $x(0)$.  While
%ideally we intend to work on the continuous-time plant, in this work
%Eq.~\eqref{eq:ode} is soundly discretized in time~\cite{fadali} into
%\begin{align}
%\label{eq:plant}
%x_{k+1} = A_d x_k+ B_d u_k
%\end{align} 
%where $k \in \mathbb N$ and $x_{0}=x(0)$ is the initial state. 
%$A_d$ and $B_d$ denote the matrices that describe the discretized plant dynamics, whereas
%$A$ and $B$ denote the continuous plant dynamics.  
%We synthesize for requirements over this discrete-time domain. 
%Later, we will address the issue of variable quantization, 
%as introduced by the ADC/DAC conversion blocks 
%(\textcolor{red}{we should include a figure that describes the control system model that we tackle}).
%\subsection{Controller synthesis via state feedback}
%\label{ssec:statefeedbackcontrol}
%Models \eqref{eq:ode} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t)$ and  $u_k$, respectively. 
%Feedback architectures can be employed to manipulate the properties and behaviors of the continuous process (the plant).   
%We are interested in the synthesis of digital feedback control algorithms, 
%as implemented on Field-Programmable Gate Arrays or Digital Signal Processors. 
%The most basic feedback architecture is the state feedback one, 
%where the control action $u_k$ (notice we work with the discretized signal) is computed by: 
%\begin{equation}
%\label{eq:controlaction}
%u_k = r_{k} - K x_k. 
%\end{equation}
%Here, $K \in \mathbb{R}^{m \times n}$ is a state-feedback gain matrix, 
%and $r_{k}$ is a reference signal (again digital).   
%The closed-loop model then takes the form 
%\begin{align}
%\label{eq:closedloopss}
%x_{k+1} = ( A_d - B_d K ) x_k + B_d r_k.
%\end{align}
%The gain matrix $K$ can be set so that the closed-loop discrete dynamics are
%shaped as desired, for instance according to a specific stability goal or
%around a specific dynamical behavior \cite{astrom1997computer}.  As argued
%later in this work, we will target more complex objectives, such as
%quantitative safety requirements, which are not typical in the digital
%control literature.  Further, we will embrace the digital nature of the
%controller, which manipulates quantized signals as discrete quantities represented with finite precision. 
%\subsection{Stability of closed-loop systems}
%\label{ssec:stability}
%In this work we employ asymptotic stability in the CEGIS loop,  
%as an objective for guessing controllers that are later proven sound over safety requirements.  
%Asymptotic stability is a property that amounts to convergence of the model executions to an equilibrium point, 
%starting from any states in a neighborhood of the point (see Figure~\ref{fig:ct} for the portrait of a stable execution, converging to the origin).  
%In the case of linear systems as in~\eqref{eq:closedloopss}, 
%considered with a zero reference signal, 
%the equilibrium point of interest is the origin. 
%A discrete-time LTI system as \eqref{eq:closedloopss} is asymptotically
%stable if all the roots of its characteristic polynomial (i.e., the
%eigenvalues of the closed-loop matrix $A_d - B_d K$) are inside the unity
%circle of the complex plane, i.e., their absolute values are strictly less than
%one~\cite{astrom1997computer} (this simple sufficient condition can be generalised, 
%however this is not necessary in our work).  
%In this paper, we express this stability specification $\phi_\mathit{stability}$ in terms of a check known as
%\emph{Jury's criterion}~\cite{fadali}: this is an easy algebraic formula to
%select the entries of matrix $K$ so that the closed-loop dynamics are shaped
%as desired.
%\subsection{Safety specifications for dynamical systems}
%\label{ssec:safety}
%We are not limited to the synthesis of digital stabilizing controllers -- a
%well known task in the literature on digital control systems -- but target
%safety requirements with an overall approach that is sound and automated. 
%More specifically, we require that the closed-loop system
%\eqref{eq:closedloopss} meets given safety specifications.  A safety
%specification gives raise to a requirement on the states of the model, such
%that the feedback controller (namely the choice of the gains matrix~$K$)
%must ensure that the state never violates the requirement.  Note that a
%stable, closed-loop system is not necessarily a safe system: indeed, the
%state values may leave the safe part of the state space while they converge
%to the equilibrium, which is typical in the case of oscillatory dynamics. 
%In~this work, the safety property is expressed as:
%\begin{equation}
%\label{eq:safetyliteral}
%\phi_\mathit{safety}\iff \forall k\ge 0.\, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}},
%\end{equation}
%where $\underline{x_{i}}$ and $\overline{x_{i}}$ are lower and upper bounds
%for the $i$-th coordinate $x_{i}$ of state $x\in \mathbb R^n$ at the $k$-th
%instant, respectively.  This means that the states will always be within an $n$-dimensional hyper-box.
%Furthermore, it is practically relevant to consider the 
%constraints $\phi_\mathit{input}$ on the input
%signal $u_{k}$ and $\phi_\mathit{init}$ on the initial states $x_0$,
%which we assume have given bounds:
%$\phi_\mathit{input} = {\forall k.\underline{u} \leq u_{k} \leq \overline{u}} $, 
%$\phi_\mathit{init} = \bigwedge_{i=1}^{n} \underline{x_{i,0}} \leq x_{i,0} \leq \overline{x_{i,0}}.$
%For the former, this means that the control input might saturates in view of
%physical constraints.
%\subsection{Performance specifications for dynamical systems}
%\label{ssec:performance}
%\textcolor{red}{we should describe here about the performance specifications for dynamical systems.}

%-------------------------------
\section{Synthesis of digital controllers with CEGIS} 
\label{sssec:cegisdig}
%-------------------------------

@@

%\textcolor{red}{[Dario: can you update the following part, aligning it with hers?]}

%  The multi-staged approach described in Sec.~\ref{sec:CEGIS-precision-incrementation}
  %uses a multi-staged verification phase and performs solution generalization based on
%the plant precision, the number of loop unrollings, 
  %and the sampling rate. Essentially, it
%  synthesizes a controller for an individual initial state
%and input with a bounded time horizon and, subsequently, it generalizes it to all reachable states,
%inputs, and time horizons during the verification phase.

The multi-staged approach above relies on the symbolic simulation over a bounded time horizon of individual initial states and inputs, 
%that form part of an uncountable space,
and on the generalization to continuous sets and over an infinite time horizon.

  %% solution over all times by refining the number of witnesses.  
  %% In order to perform the safety check, this approach uses abstract acceleration
  %% \cite{cattaruzza2015unbounded} to compute the controller's set of states. 
  %% \todo{clarify: what is the ``set of states'' of the controller? }
%\end{itemize}

%% The verification of a hybrid continuous/discrete system over an
%% unbounded time, requires the non-deterministic simulation of an
%% infinite set of states, which is undecidable.

%% Our multi-staged approach
%% creates an abstraction of the system by establishing bounds on the
%% error between such a system and a discrete-time/continuous-space
%% model.  This model is still however, quite complex, hard to verify,
%% and many of the states involved are inconsequential to the safety
%% specification.

@@

The CEGIS loop for this approach is illustrated in Figure~\ref{fig:CEGARIS}. 
%
% Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
% Our model, rather than look at a bisimulation \todo{specific notion that has not been formally define - do not use} in discrete time, explores a much higher abstraction, 
% evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specification.
% We have also identified 
% the quantization noise as a bounded non-deterministic input.
It comprises the following steps:  
\begin{enumerate}
\item 
  We start by doing some preprocessing:%%  that 
  %% can be computationally intensive and we don't want it to be part of the CEGIS loop.
% Let us recall the formulas used for single-input and single-output (SISO) systems 
  % in sections \ref{sec:reachable} and \ref{sec:observable}.
  \begin{enumerate}
\item Compute the characteristic polynomial of the matrix $(A_d-B_dK)$ as 
$P_a(z) = z^n+\sum_{i=1}^n{(a_i-k_i)z^{n-i}}$.  As with the multi-staged approach, this polynomial must satisfy Jury's criterion.
\reply{The polynomial $P_a(z)$ is obtained by transforming the system into controllable form( See~\cite{astrom1997computer}, where both the dynamics and the controller are multiplied by a matrix $\mat{T}$ (derived from $\mat{A}$ and $\mat{B}$), which is given to the synthesiser.  The synthesizer is thus able to provide a $\mat{K}$ such that $k_i$ are the elements  of $\mat{K}\mat{T}^{-1}$.
}
%\blue{[shouldn't there be a b term in the polynomial's coefficients here? No. the polynomial only evaluates the poles]}
%%   A controllable system with quantization will have the following dynamics, where
%% we use underscript ``n'' to denote a noise term,
%% ``cf'' to denote Controllable Canonical Form,
%% $\nu_1$, $\nu_2$ are quantization errors
%% caused by the ADC and DAC conversions, respectively, 
%% and $\nu_3$ is the round-off error.
%% \todo{Why do we use $_t$ here?}
%% %
%% \begin{align}
%% \label{eq:observer_LTI_cf}
%% x_{k+1}=&\mat{A}_{t}x_k+\mat{B}_{t} r_k+\mat{B}_{n} (\nu_1 + \nu_2 + \nu_3)
%% \end{align}
%% \begin{align*}
%% \mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
%% \mat{B}_{t}=&\mat{T}\mat{B}_d&
%% \mat{B}_{n}=&[1 \cdots 1]^T&
%% \mat{T}=&\mat{W}_{cf}\mat{W}^{-1}&
%% \end{align*}

%% \begin{align*}
%% \mat{A}_{t}=\mat{T}\mat{A}_d\mat{T}^{-1}=&\left[
%% \begin{array}{ccccc}
%% 0&1&0&\cdots&0\\
%% 0&0&1&\cdots&0\\
%% \vdots&\vdots&\vdots&\ddots&\vdots\\
%% 0&0&0&\cdots&1\\
%% k_n-a_n&k_{n-1}-a_{n-1}&k_{n-2}-a_{n-2}&\cdots&k_1-a_1
%% \end{array}\right]
%% \label{eq:cf_SISO_2}
%% \end{align*}

%% \begin{enumerate}
%% \item Calculate $\mat{T}$ as $\mat{W}_{cf}\mat{W}^{-1}$. %using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}
%% \item Extract the characteristic polynomial of the above matrix:
%% %
%\begin{align*}
%$P_a = z^n+\sum_{i=1}^p{(a_i-k_i)z^{n-i}}$
%\end{align*}

@@

\end{enumerate}
\item In the {\sc synthesize} phase, we synthesize a candidate controller
%We begin with a minimum set of linear inequalities describing the constraints caused by the input and the initial state during a single iteration, in addition to a stability 
% specification defined by Jury's 
%criteria. %, and try to synthesise a controller that meets these specs.
% For a stabilized closed-loop, which is necessary for our safety property, we need
% these roots to remain within the unit circle, which we verify using Jury's criteria.
% Notice that in our polynomials $c_0=1$ and we only use reals.
% Once we have established convergence, we may examine the 
% safety specification $\phi_\mathit{safety}$.
  $K$ 
  that satisfies
  $\phi_\mathit{stability} \wedge \phi (K)$ by invoking a SAT solver.  
  The safety specification given to the SAT formula is an over-approximation of the original safety specification that only has constraints for a given set of iterations ($k$) and initial values.  
 The program search algorithm is as in the multi-staged approach.  
%%   We obtain it by solving the SAT formula 
%% $$\mathcal{J}_K=\mathcal{J}\langle I,F \rangle (P_{a-k},\mat{T})\wedge \phi_\mathit{safety}$$ 
%% that satisfies the input constraint and Jury's criteria for 
%% $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \tilde{\mat{K}}=\mat{K}\mat{T}$,
%% where $\mathcal{J}\langle I,F \rangle (P,\mat{T})$ is a program describing Jury's method.
If there is no candidate solution we return UNSAT and exit the loop. 
\item Once we have a candidate solution, we perform a safety verification % If this step fails, we find a counterexample iteration and initial state and create a new constraint to refine our abstraction.
  %
  of the 
  %consisting of evaluating the
  progression of the system from $\phi_\mathit{init}$ over time,
$x_{k} \models \phi_\mathit{safety}$. %\blue{why $k+1$ here?}. %% For this purpose we require 
%% an initial set $\phi_\mathit{init}= x_0 \in [\underline{x_0} \ \overline{x_0}]$ 
%% from which the system progresses.
% Obviously, this set $x_0$ 
% must satisfy the specification; otherwise, the system will be unsafe to begin with.
% We accept a specification in the form $\mat{E}\mat{T}x_0<\mat{f}$. 
% The presence of $\mat{T}$ is because this will typically be given in the original 
% state-space.
%% Next, we use abstract acceleration to get a second abstraction that 
  %% encompasses the accelerated continuous reach-tube over an unbounded time.
  In order to compute the progression of point $x_0$ at iteration $k$,
  we consider the solution of the closed-loop model 
  %we take the
  %the closed-loop model:
  %$x_{k+1}=\mat{A}_{t}x_k+\mat{B}_{n} (\nu_1 + \nu_2 + \nu_3)$
  %and accelerate it to:
%
{
\scriptsize
\begin{align*}
%\begin{split}
x=(A_d-B_dK)^kx_0+ \sum_{i=0}^{k-1} (A_d-B_dK)^i B_{n}(\nu_1+\nu_2+\nu_3), 
%\end{split}
\end{align*}
}
where $B_n= [1 \cdots 1]^T$. 
%\label{eq:acc_observer_LTI_cf}
%+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{t} r_i 
%
As we require to verify the closed-loop model for any $k$ up to infinity, 
we use Abstract Acceleration again to compute the reach-tube, 
i.e., the set of all reachable states at all times given an initial set
$\phi_\mathit{init}$:
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=\mathcal{A} X_0 + \mathcal{B}_{n} N, \,
X_0 =\left \{x: x \models \phi_\mathit{init} \right\}, 
\end{align} 
%
where $\mathcal{A}=\bigcup_{k=1}^\infty (A_d-B_dK)^k,
\mathcal{B}_{n}=\bigcup_{k=1}^\infty \sum_{i=0}^k(A_d-B_dK)^iB_{n}$ are
abstract matrices for the closed-loop system~\cite{cattaruzza2015unbounded},
whereas the set $N$ is obtained above.

@@

%-------------------------------
%\subsection{Formal specification of transient response} 
%\label{ssec:transientspecification}
%-------------------------------

%In a SISO system, there are 2 values that typically define the response-time of the system (using a step response).
%\begin{enumerate}
%\item The \emph{rise time} is defined as the time it takes the output to reach a percentage (eg $90\%$) of the target value given a step chnage in the input.
%\item The \emph{settling time} is defined as the time it takes the output to stabilize within a margin of the target value (eg $10\%$).
%\end{enumerate} 
%In the case of first and second order systems (with ), the decay is governed by the equations 
%\begin{align*}
%&y_{decay}=
%&\left\{
%\begin{array}{cc}
%e^{-|\omega| t}& 1^{st} order\\
%e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
%e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
%\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
%\end{align*}
%where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
%These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
%All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.
%\begin{theorem}
%Given a set of eigenvalues for a discrete time model of an $n^{th}$ order system $\lambda_i =e^{\omega_i}: i \in [1\ n]$, the settling time of the system is upperbounded by
%\begin{equation}
%t_s<-log_{|\overline{\lambda}|}({p_s}) : |\overline{\lambda}| =\max(|\lambda_i|) \wedge |\overline{\lambda}|<1,\ \forall i \in [1\ n]
%\label{eq:set_time}
%\end{equation}
%where $0\leq p_s \leq 1$ is the desired percentage of the target within to remain after $t_s$.
%\end{theorem} 
%\begin{proof}
%Given a closed loop single output LTI system of order $n$ with dynamics
%\begin{equation}
%\left\{\begin{array}{c}\vec{x}_k=\mat{A}\vec{x}_{k-1}\\ \vec{y}_k=\mat{C}\vec{x}_k\end{array}\right. \Rightarrow \left\{ \begin{array}{c}\vec{x}_k'=\mat{J}\vec{x}_{k-1}'\\ \vec{y}_k=\mat{C}\mat{S}\vec{x}_k'\end{array}\right. : \left\{ \begin{array}{c}\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \\ \vec{x}_k'=\mat{S}^{-1}\vec{x}_k\end{array}\right.
%\label{eq:eigendynamics}
%\end{equation}
%
%the output decay is a linear combination of the decay of eigenvalues or
%eigenvalue pairs.  
%Let $\mat{C}'\in \mathbb{R}^{1 \times n}=\mat{C}\mat{S}$. Since $\mat{C}'$ is constant, the output $y_k$ is a linear combination of the states $\vec{x}'_k$. We accelerate equation \eqref{eq:eigendynamics} to obtain
%\begin{align}
%y_k&=\mat{C}'\mat{J}^k\vec{x}_0 = \sum_{i=0}^n \left( c_i \lambda_i^k + c_{il}\binom{k}{l} \lambda_i^{k-l}\right)\\
%|y_k|&\leq \sum_{i=0}^n \left|c_i + c_{il}\binom{k}{l} \lambda_i^{-l}\right||\lambda_i^k|
%\end{align}
%where the second  part of the equation corresponds to the upper diagonal terms of the Jordan blocks. We also have $|\lambda_i^k|\leq |\overline{\omega}^k|$,hence
%\begin{align}
%|y_k|&\leq |\overline{\omega}^k|\sum_{i=0}^n \left|c_i + c_{il}\binom{k}{l} \lambda_i^{-l}\right|
%\end{align}
%Since we are synthesising the eigenvalues, rather than solve this formula for complex Jordan Shapes, we will restrict our synthesiser to provide solutions with distinct eigenvalues (which is by far the most common case). Therefore the right hand terms dissappear and we end up with
%\begin{align}
%|y_k|&\leq |\overline{\omega}^k|\sum_{i=0}^n |c_i|
%\end{align}

>>>>

The time discretization in Sec.~\ref{sec:model} only allows us to evaluate
the state of the plant at integer multiples of the sample time.  
%\note{Unlike the CAV18 paper, these gain matrices have not been discussed yet. }
However, in order to synthesize a soundly safe controller, 
% $\mat{K}$, 
we must be able to ensure that the safety of the plant holds at any continuous point in time.  
In order to do so, we accelerate the dynamics of the plant according to the continuous-time model
semantics in Sec.~\ref{sec:model}.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Accelerated\! dynamics\! for\! continuous-time\! models\! with\! parametric\! inputs}\label{sec:real_discrete_param_inputs}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
% 
Given a specific time $T$, the state $\vec{x}(t=T)$ can be expressed as a
linear function of initial state $\vec{x}_0$ and parametric input $\vec{u}$ as follows.

@@

\fi
%-------------------------------
\section{Synthesis of digital controllers with CEGIS} 
\label{sssec:cegisdig}
%-------------------------------

@@

The multi-staged approach above relies on the symbolic simulation over a bounded time horizon of individual initial states and inputs, 
and on the generalization to continuous sets and over an infinite time horizon.

@@

The CEGIS loop for this approach is illustrated in Figure~\ref{fig:CEGARIS}. 
%

It comprises the following steps:  
\begin{enumerate}
\item 
  We start by doing some preprocessing:%%  that 
  %% can be computationally intensive and we don't want it to be part of the CEGIS loop.
% Let us recall the formulas used for single-input and single-output (SISO) systems 
  % in sections \ref{sec:reachable} and \ref{sec:observable}.
  \begin{enumerate}
\item Compute the characteristic polynomial of the matrix $(A_d-B_dK)$ as 
$P_a(z) = z^n+\sum_{i=1}^n{(a_i-k_i)z^{n-i}}$.  As with the multi-staged approach, this polynomial must satisfy Jury's criterion.
\reply{The polynomial $P_a(z)$ is obtained by transforming the system into controllable form( See~\cite{astrom1997computer}, where both the dynamics and the controller are multiplied by a matrix $\mat{T}$ (derived from $\mat{A}$ and $\mat{B}$), which is given to the synthesiser.  The synthesizer is thus able to provide a $\mat{K}$ such that $k_i$ are the elements  of $\mat{K}\mat{T}^{-1}$.
}

@@

\end{enumerate}
\item In the {\sc synthesize} phase, we synthesize a candidate controller
  $K$ 
  that satisfies
  $\phi_\mathit{stability} \wedge \phi (K)$ by invoking a SAT solver.  
  The safety specification given to the SAT formula is an over-approximation of the original safety specification that only has constraints for a given set of iterations ($k$) and initial values.  
 The program search algorithm is as in the multi-staged approach.  
If there is no candidate solution we return UNSAT and exit the loop. 
\item Once we have a candidate solution, we perform a safety verification % If this step fails, we find a counterexample iteration and initial state and create a new constraint to refine our abstraction.
  %
  of the 
  %consisting of evaluating the
  progression of the system from $\phi_\mathit{init}$ over time,
$x_{k} \models \phi_\mathit{safety}$. 
  In order to compute the progression of point $x_0$ at iteration $k$,
  we consider the solution of the closed-loop model 
  %we take the
  %the closed-loop model:
  %$x_{k+1}=\mat{A}_{t}x_k+\mat{B}_{n} (\nu_1 + \nu_2 + \nu_3)$
  %and accelerate it to:
%
{
\scriptsize
\begin{align*}
%\begin{split}
x=(A_d-B_dK)^kx_0+ \sum_{i=0}^{k-1} (A_d-B_dK)^i B_{n}(\nu_1+\nu_2+\nu_3), 
%\end{split}
\end{align*}
}
where $B_n= [1 \cdots 1]^T$. 
%\label{eq:acc_observer_LTI_cf}
%+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{t} r_i 
%
As we require to verify the closed-loop model for any $k$ up to infinity, 
we use Abstract Acceleration again to compute the reach-tube, 
i.e., the set of all reachable states at all times given an initial set
$\phi_\mathit{init}$:
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=\mathcal{A} X_0 + \mathcal{B}_{n} N, \,
X_0 =\left \{x: x \models \phi_\mathit{init} \right\}, 
\end{align} 
%
where $\mathcal{A}=\bigcup_{k=1}^\infty (A_d-B_dK)^k,
\mathcal{B}_{n}=\bigcup_{k=1}^\infty \sum_{i=0}^k(A_d-B_dK)^iB_{n}$ are
abstract matrices for the closed-loop system~\cite{cattaruzza2015unbounded},
whereas the set $N$ is obtained above.

@@
